---
title: "Analyze the Popularity of Programming Languages"
output:
  html_document:
    df_print: paged
---

# Description 


Project prompt: 

> How can you determine which programming languages and technologies are most widely used? Which languages are gaining or losing popularity, helping you decide where to focus your efforts?
>
> One excellent data source is Stack Overflow, a programming question-and-answer site with more than 16 million questions on programming topics. Each Stack Overflow question is tagged with a label identifying its topic or technology. By counting the number of questions related to each technology, you can estimate the popularity of different programming languages.
> 
> In this project, you will use data from the [Stack Exchange Data Explorer](https://data.stackexchange.com/) to examine the relative popularity of R compared to other programming languages.
> 
> You'll work with a dataset containing one observation per tag per year, including the number of questions for that tag and the total number of questions that year.

`stack_overflow_data.csv`

|Column|Description|
|------|-----------|
|`year`|The year the question was asked (2008-2020)|
|`tag`|A word or phrase that describes the topic of the question, such as the programming language|
|`num_questions`|The number of questions with a certain tag in that year|
|`year_total`|The total number of questions asked in that year|


# Assumptions


Assumptions about dataset:  

- All questions about R are tagged with "r".  
- Each combination of year and tag occurs no more than once. (i.e. tags are never duplicated within the same year)


# Setup


```{r}
# Load necessary packages
library(readr)
library(dplyr)
library(ggplot2)
library(stringr)
```

```{r}
# Load the dataset
data <- read_csv("stack_overflow_data.csv")
```
```{r}
# View the dataset
head(data)
```


# Analysis 


## R-related questions 


Question 1: 

> What was the percentage of R questions in 2020? Save the result in a data frame, r_2020, containing five columns: year, tag, num_questions, year_total, and percentage.

```{r}
r_2020 <- data %>% 
	filter(year == 2020, tag == "r") %>%
	mutate(percentage = num_questions / year_total * 100)

r_2020
```


## Most-asked-about programming languages

Question 2: 

> Identify the five programming language tags with the highest total number of questions asked between 2015 and 2020 (inclusive). Save the tag names as "highest_tags". This variable can be a character vector, tibble, or data frame (if the latter, please use the column name "tag"). 

```{r}
highest_tags <- data %>%
	# Filter for language tags with questions asked between 2015 and 2020
	filter(year >= 2015 & year <= 2020) %>%
	# For each programming language tag, calculate total number of questions asked 
	group_by(tag) %>%
	summarise(total_num_questions = sum(num_questions)) %>%
	ungroup() %>%
	# Then get the five language tags with highest number of questions asked
	slice_max(total_num_questions, n=5) %>%
	arrange(desc(total_num_questions))

highest_tags
```

```{r}
# Visualise this data
ggplot(highest_tags, aes(x = reorder(tag, -total_num_questions), y = total_num_questions, fill = tag)) + 
	geom_col() + 
	labs(x = "Programming language", 
		 y = "Number of questions asked", 
		 title = str_wrap("Programming Languages with the Most Questions Asked on Stack Overflow Between 2015 
						  and 2020", 60),
		 alt = "A bar plot showing the number of questions asked about five programming languages")
```


# Final thoughts and further questions


I would like to look further into the context in which these languages have become popular-- some questions might include:  

- What popular technologies were introduced between 2008-2020 that are related to the five programming language tags with the most questions? 
  - For example, React was released in 2013 and is maintained and used by Facebook-- how did this affect the number of JavaScript-related questions?
  - Or, did the number of Python questions at the same time as, or shortly after, the release of ChatGPT in 2022?) 
